---
title: "HW 02 - Data Wrangling"
output: html_document
author: Hanke Zheng
link-citations: yes
---
### Download and Read in the data

First download and then read in with data.table:fread()

```{r, echo=TRUE, message=FALSE}
# download individual/regional data from github
download.file("https://raw.githubusercontent.com/USCbiostats/data-science-data/master/01_chs/chs_individual.csv", "chs_individual.csv", method="libcurl", timeout = 60)
individual <- data.table::fread("chs_individual.csv")

download.file("https://raw.githubusercontent.com/USCbiostats/data-science-data/master/01_chs/chs_regional.csv", "chs_regional.csv", method="libcurl", timeout = 60)
regional <- data.table::fread("chs_regional.csv")

library(data.table)
library(dtplyr)
library(dplyr)

```

There are 1200 raws and 23 coloumns in the individual dataset; 12 raws and 27 variables in the regional dataset. 
### Merge
```{r}
## merge the two dataset by location
ind_region <- merge(x = individual, y = regional, by.x = "townname", by.y = "townname", all.x = TRUE, all.y = FALSE) 

count_n <- merge(x = individual, y = regional, by.x = "townname", by.y = "townname", all.x = TRUE, all.y = FALSE) %>%nrow()

```
The # of the merged dataset is the same as that of the individual dataset - no duplicates. 

### Manipulate the data
```{r}
# check the missing values for bmi
message("Missing: ", ind_region[is.na(bmi), .N])
# impute data using the average within the variables “male” and “hispanic"
ind_region[, bmi_imp := fcoalesce(bmi, mean(bmi, na.rm = TRUE)),
    by = .(male, hispanic)]
ind_region$bmi[is.na(ind_region$bmi)] <- ind_region$bmi_imp

# check the missing values for fev
message("Missing: ", ind_region[is.na(fev), .N])
# impute data using the average within the variables “male” and “hispanic"
ind_region[, fev_imp := fcoalesce(fev, mean(fev, na.rm = TRUE)),
    by = .(male, hispanic)]
ind_region$fev[is.na(ind_region$fev)] <- ind_region$fev_imp

# Create a new categorical variable named “obesity_level” using the BMI measurement (underweight BMI<14; normal BMI 14-22; overweight BMI 22-24; obese BMI>24). 
ind_region$obesity_level <- ifelse(ind_region$bmi>= 14 & ind_region$bmi<= 22, "normal", 
                         ifelse(ind_region$bmi<14, "underweight",
                                ifelse(ind_region$bmi > 24, "obese", "overweight")))
table(ind_region$obesity_level)
range(ind_region$bmi)


# check the missing values for binary variables 'smoke' and 'gasstove'
message("Missing: ", ind_region[is.na(smoke), .N])
message("Missing: ", ind_region[is.na(gasstove), .N])

# Create another categorical variable named “smoke_gas_exposure” that summarizes “Second Hand Smoke” and “Gas Stove.” The variable should have four categories in total.
ind_region$smoke_gas_exposure <- ifelse(ind_region$smoke== 1 & ind_region$gasstove == 1, "exposed to both", 
                         ifelse(ind_region$smoke==1 & ind_region$gasstove != 1, "exposed to smoke",
                        ifelse(ind_region$smoke !=1 & ind_region$gasstove == 1, "exposed to gas stove", "not exposed to both")))
table(ind_region$smoke_gas_exposure)

```

# Summary Table
```{r}
#Create four summary tables showing the average (or proportion, if binary) and sd of “Forced expiratory volume in 1 second (ml)” and asthma indicator by town, sex, obesity level, and “smoke_gas_exposure.”
# convert to data frame
ind_region_df <- as.data.frame(ind_region)
# by town
ind_region_df %>%
  group_by(townname) %>%
  summarise(fev_ave = mean(fev),
            fev_sd = sd(fev),
            prop_asthma = mean(asthma, na.rm = TRUE),
            asthma_sd = sd(asthma, na.rm = TRUE)
  )

# by sex
ind_region_df %>%
  group_by(male) %>%
  summarise(fev_ave = mean(fev),
            fev_sd = sd(fev),
            prop_asthma = mean(asthma, na.rm = TRUE),
            asthma_sd = sd(asthma, na.rm = TRUE)
  )

# by obesity level
ind_region_df %>%
  group_by(obesity_level) %>%
  summarise(fev_ave = mean(fev),
            fev_sd = sd(fev),
            prop_asthma = mean(asthma, na.rm = TRUE),
            asthma_sd = sd(asthma, na.rm = TRUE)
  )


# by smoke_gas_exposure
ind_region_df %>%
  group_by(smoke_gas_exposure) %>%
  summarise(fev_ave = mean(fev),
            fev_sd = sd(fev),
            prop_asthma = mean(asthma, na.rm = TRUE),
            asthma_sd = sd(asthma, na.rm = TRUE)
  )



```

### 4. Exploratory Data Analysis
- 1. What is the association between BMI and FEV (forced expiratory volume)? 
- 2. What is the association between smoke and gas exposure and FEV? 
- 3. What is the association between PM2.5 exposure and FEV?


# Facet plot showing scatterplots with regression lines of BMI vs FEV by “townname”.
```{r}
library(leaflet)
library(tidyverse)
library(ggplot2)

ind_region_df %>%
  filter(!(townname %in% NA)) %>%
  ggplot(mapping=aes(x=bmi, y=fev, color=townname)) +
  geom_point()+
  stat_smooth(method=lm)+
  facet_wrap(~townname)
  
```
A positive association was observed between BMI and FEV based on the scatterplots.


#Stacked histograms of FEV by BMI category and FEV by smoke/gas exposure. Use different color schemes than the ggplot default.
```{r}
ind_region_df %>%
  ggplot()+
  geom_histogram(mapping=aes(x=fev,fill=obesity_level))+
  # change the default palette
  scale_fill_brewer(palette = "PuOr")+
  labs(title="FEV by BMI category and by obesity level", x="FEV", y="count")
  theme()


ind_region_df %>%
  filter(!(smoke_gas_exposure %in% NA)) %>%
  ggplot()+
  geom_histogram(mapping=aes(x=fev,fill=smoke_gas_exposure))+
  # change the default palette
  scale_fill_brewer(palette = "Zissou")+
  labs(title="FEV by BMI category and by smoke/gas exposure", x="FEV", y="count")
  theme() 
  
  
```
FEV is normally distributed by obesity level and exposure status. 


# Barchart of BMI by smoke/gas exposure.
```{r}
ind_region_df %>%
  filter(!(smoke_gas_exposure %in% NA)) %>%
  ggplot()+
  geom_bar(mapping=aes(x=smoke_gas_exposure,fill=obesity_level))+
  # change the default palette
  scale_fill_brewer(palette = "PuOr")+
  labs(title="BMI level by smoke/gas exposure", x="Exposure Status", y="count")
  theme()

```
The majority of people in different exposure status are within normal BMI level.

#Statistical summary graphs of FEV by BMI and FEV by smoke/gas exposure category.
```{r}

ind_region_df %>%
ggplot() +
  geom_boxplot(mapping=aes(y=fev, fill=obesity_level))   



ind_region_df %>%
    filter(!(smoke_gas_exposure %in% NA)) %>%
    ggplot() +
    geom_boxplot(mapping=aes(y=fev, fill=smoke_gas_exposure))   
```


#A leaflet map showing the concentrations of PM2.5 mass in each of the CHS communities.

```{r}
library(leaflet)
leaflet(ind_region_df) %>%
  addProviderTiles('OpenStreetMap') %>%
  addCircles(lat=~lat, lng=~lon, color="green",opacity=1,
             fillOpacity=1, radius=~(pm25_mass*300))
```

The PM2.5 level is relatively high in LA areas in comparative to areas in the northern California.

# Choose a visualization to examine whether PM2.5 mass is associated with FEV.

```{r}
ind_region[,fev_ave :=mean(fev),by=townname]
ind_region %>%
  ggplot(mapping = aes(x=pm25_mass, y=fev_ave))+
  geom_point()+
  stat_smooth(method=lm)+
  labs(titles="FEV VS PM2.5 mass", x="PM2.5 mass", y="FEV")

```

Using the average FEV for each town, a negatvie association between PM2.5 and FEV was observed. 





