---
title: "Assignment 04 "
output: github_document
author: Hanke Zheng
---

# HPC
## Problem 1: Make sure the code is nice
Rewrite the R functions to make them faster (take a look at stackoverflow and google).

```{r}
# Total row sums
fun1 <- function(mat) {
  n <- nrow(mat)
  ans <- double(n) 
  for (i in 1:n) {
    ans[i] <- sum(mat[i, ])
  }
  ans
}

fun1alt <- function(mat) {
 rowSums(mat, na.rm = FALSE)
}

# Cumulative sum by row
fun2 <- function(mat) {
  n <- nrow(mat)
  k <- ncol(mat)
  ans <- mat
  for (i in 1:n) {
    for (j in 2:k) {
      ans[i,j] <- mat[i, j] + ans[i, j - 1]
    }
  }
  ans
}

fun2alt <- function(mat) {
  ans <- mat
  for (i in 1:nrow(mat)) {
    ans[i, ] <- cumsum(mat[i, ])
  }
  ans}

# Use the data with this code
set.seed(2315)
dat <- matrix(rnorm(200 * 100), nrow = 200)

# Test for the first
microbenchmark::microbenchmark(
  fun1(dat),
  fun1alt(dat), unit = "relative", check = "equivalent"
)

# Test for the second
microbenchmark::microbenchmark(
  fun2(dat),
  fun2alt(dat), unit = "relative", check = "equivalent"
)
```
## Problem 2: Parallel computing
The following function allows simulating PI
```{r}
sim_pi <- function(n = 1000, i = NULL) {
  p <- matrix(runif(n*2), ncol = 2)
  mean(rowSums(p^2) < 1) * 4
}

# Here is an example of the run
set.seed(156)
sim_pi(1000) # 3.132
```

In order to get accurate estimates, we can run this function multiple times, with the following code:
```{r}
# This runs the simulation a 4,000 times, each with 10,000 points
set.seed(1231)
system.time({
  ans <- unlist(lapply(1:4000, sim_pi, n = 10000))
  print(mean(ans))
})
```


Rewrite the previous code using parLapply() to make it run faster. Make sure you set the seed using clusterSetRNGStream():
```{r}
# YOUR CODE HERE
library(parallel)
system.time({
  cl <- makePSOCKcluster(2L, setup_strategy = "sequential") # STEP 1: Make cluster
  clusterSetRNGStream(cl, 1231) # Set seed
  ans <- unlist(parLapply(cl,1:4000, sim_pi, n = 10000))
  print(mean(ans))
  stopCluster(cl)
})
```

# SQL
## Set up a temporary database
```{r}
#install.packages(c("RSQLite", "DBI"))
library(RSQLite)
library(DBI)

# Initialize a temporary in memory database
con <- dbConnect(SQLite(), ":memory:")

# Download tables
film <- read.csv("https://raw.githubusercontent.com/ivanceras/sakila/master/csv-sakila-db/film.csv")
film_category <- read.csv("https://raw.githubusercontent.com/ivanceras/sakila/master/csv-sakila-db/film_category.csv")
category <- read.csv("https://raw.githubusercontent.com/ivanceras/sakila/master/csv-sakila-db/category.csv")

# Copy data.frames to database
dbWriteTable(con, "film", film)
dbWriteTable(con, "film_category", film_category)
dbWriteTable(con, "category", category)

```


## Question 1: How many movies is there available in each rating category? 
```{sql, connection=con}
SELECT rating, COUNT(*) AS count
FROM film
GROUP BY rating
```
There are 180, 210, 194, 223, and 195 films grouped in G, NC-17, PG, PG-13, R.

## Question 2: What is the average replacement cost and rental rate for each rating category.
SELECT avg(amount), min(amount), max(amount), sum(amount)
FROM payment
```{sql, connection=con}
SELECT avg(replacement_cost),avg(rental_rate),rating
FROM film
GROUP BY rating
```
In rating group G, NC-17, PG, PG-13, and R, the average replacement costs are 20.12, 20.14, 18.96, 20.4, and 20.23; the average rental rate are 2.91, 2.97, 3.05, 3.03, and 2.94.

## Question 3: Use table film_category together with film to find the how many films there are with each category ID

```{sql, connection=con}
SELECT a.film_id, count(b.film_id) as COUNT,a.category_id, b.film_id
FROM film_category AS a
  INNER JOIN film AS b on a.film_id = b.film_id
GROUP BY category_id
ORDER BY COUNT DESC 
```
Films with category ID of 10 are them are the most popular ones.

## Question 4: Incorporate table category into the answer to the previous question to find the name of the most popular category.


```{sql, connection=con}
SELECT name, a.category_id , count(a.film_id) as COUNT
FROM film_category a 
  JOIN film b on a.film_id=b.film_id
  JOIN category c on a.category_id=c.category_id 
GROUP BY a.category_id
ORDER BY COUNT DESC
```

The most popular category is Sprorts followed by Foreign, with 74 films in Sports and 73 in Foreign.

