---
title: "Lab 04 - Data Visualization"
output: html_document
author: Hanke Zheng
link-citations: yes
---



```{r setup, message=FALSE, echo=FALSE, warning=FALSE}
#install.packages(c("data.table","leaflet"))
library(data.table)
library(leaflet)
library(tidyverse)
```

# Learning Goals

- Read in and prepare the meteorological dataset
- Create several graphs with different `geoms()` in `ggplot2`
- Create a facet graph
- Conduct some customizations to the graphs
- Create a more detailed map using `leaflet()`


# Lab Description

We will again work with the meteorological data presented in lecture.

**The objective of the lab is to examine the association between monthly average dew point temperature and relative humidity in four regions of the US and by elevation.**

# Steps

### 1. Read in the data

First download and then read in with data.table:fread()

```{r, echo=TRUE, message=FALSE}
download.file("https://raw.githubusercontent.com/USCbiostats/data-science-data/master/02_met/met_all.gz", "met_all.gz", method="libcurl", timeout = 60)
met <- data.table::fread("met_all.gz")
```

### 2. Prepare the data

- Remove temperatures less than -17C
- Make sure there are no missing data in the key variables coded as 9999, 999, etc
- Take monthly averages by weather station
- Create a region variable for NW, SW, NE, SE based on lon = -98.00 and lat = 39.71 degrees
- Create a categorical variable for elevation as in the lecture slides

```{r}

met<-met[met$temp > -17]
met[met$elev==9999.0] <- NA

#na.rm=TRUE: remove the NA for calculation 
met_avg<-met[,.(temp=mean(temp,na.rm=TRUE), rh=mean(rh,na.rm=TRUE),wind.sp=mean(wind.sp,na.rm=TRUE), dew.point=mean(dew.point, na.rm=TRUE), vis.dist=mean(vis.dist,na.rm=TRUE), lat=mean(lat), lon=mean(lon),
elev=mean(elev,na.rm=TRUE)), by=c("USAFID")]

met_avg$elev_cat <- ifelse(met_avg$elev> 252, "high", "low")

met_avg$region <- ifelse(met_avg$lon> -98 & met_avg$lat > 39.71, "north east", 
                         ifelse(met_avg$lon> -98 & met_avg$lat < 39.71, "south east",
                                ifelse(met_avg$lon< -98 & met_avg$lat > 39.71, "north west", "south west")))
table(met_avg$region)

# met_avg$region <- ifelse(met_avg$lon> -98, "east", "west")


```

### 3. Use `geom_boxplot` to examine the dew point temperature and relative humidity by region

- Use facets
- Make sure to deal with `NA` categoryan
- Describe what you observe in the graph
```{r}
met_avg %>%
  filter(!(region %in% NA)) %>%
ggplot()+
  geom_boxplot(mapping=aes(x=region, y=dew.point))


met_avg %>%
  filter(!(region %in% NA)) %>%
ggplot() +
  geom_boxplot(mapping=aes(y=dew.point, fill=region)) + 
  facet_wrap(~region,nrow=2)


met_avg %>%
  filter(!(region %in% NA)) %>%
ggplot() +
  geom_boxplot(mapping=aes(y=rh, fill=region)) + 
  facet_wrap(~region,nrow=2)
# 2 by 2
```
- Dew point tem is higher in the south east region
- Relative humidity is higher in south east region

### 4. Use `geom_point` with `stat_smooth` to examine the association between dew point temperature and relative humidity by region

- Colour points by region
- Make sure to deal with `NA` category
- Fit a linear regression line by region
- Describe what you observe in the graph

```{r}
met_avg %>%
  filter(!(region %in% NA)) %>%
  ggplot(mapping=aes(x=dew.point, y=rh, color=region))+
  geom_point()+
  stat_smooth(method=lm)

  # stat_smooth: linear regression line
```
- The association between dew point temp and relative humidity is positive in all region.
- The slope is greater in north west. 
### 5. Use `geom_bar` to create barplots of the weather stations by elevation category coloured by region

- Bars by elevation category
- Change colours from the default. Colour by region using `scale_fill_brewer` see [this](http://rstudio-pubs-static.s3.amazonaws.com/5312_98fc1aba2d5740dd849a5ab797cc2c8d.html)
- Create nice labels on axes and add a title
- Describe what you observe in the graph
- Make sure to deal with NAs
```{r warning=FALSE, message=FALSE}
met_avg %>%
  filter(!(region %in% NA)) %>%
  ggplot()+
  geom_bar(mapping=aes(x=elev_cat,fill=region))+
  # change the default palette
  scale_fill_brewer(palette = "PuOr")+
  labs(title="Number of weather stations by levation category and region", x="Elevation Category", y="Count")
  theme()
  
```

### 6. Use `stat_summary` to examine mean dew point and relative humidity by region with standard deviation error bars

- Make sure to remove `NA`
- Use fun.data="mean_sdl" in `stat_summary`
- Describe the graph and what you observe

```{r}
met_avg %>%
  filter(!(region %in% NA)) %>%
  ggplot(mapping = aes(x=region,y=dew.point))+
  stat_summary(fun.data = "mean_sdl")
  

met_avg %>%
  filter(!(region %in% NA)) %>%
  ggplot(mapping = aes(x=region,y=rh))+
  stat_summary(fun.data = "mean_sdl")
```


- Dew point temperature is the highest and concentrated in south east whereas it's the lowest and more spread-out in north west.
- Relative humidity is higher and more concentrated in north east and south east while it's dryer in south west and north west.


### 7. Make a map showing the spatial trend in relative humidity in the US

- Make sure to remove `NA`
- Use leaflet()
- Make a colour palette with custom colours
- Add a legend

```{r}
met_avg2 <- met_avg [!is.na(rh)]
rh_pal=colorNumeric(c('blue', 'purple', 'red'), domain=met_avg2$rh)

leaflet(met_avg2) %>%
  addProviderTiles('OpenStreetMap') %>%
  addCircles(lat=~lat, lng=~lon, color=~rh_pal(rh),label=~paste0(round(rh,2),'rh'), opacity = 1,fillOpacity = 1, radius=500) %>%
  addLegend('bottomleft', pal=rh_pal, values = met_avg$rh, title="Relative Humidity", opacity = 1)


```

- The relative humidity is increasing from west to east. 

### 8. Use a ggplot extension

- Pick and extension (except cowplot) from [here](https://exts.ggplot2.tidyverse.org/gallery/) and make a plot of your choice using the met data (or met_avg)
- Might want to try examples that come with the extension first (e.g. ggtech, gganimate, ggforce)

```{r}
library(ggstance)
met_avg %>%
  filter(!(region %in% NA)) %>%
  ggplot(mapping = aes(y=region,x=dew.point, fill=elev_cat))+
  geom_boxploth()

###write.csv(met_avg,"path/.csv")
###saveRDS(met_avg,"path/.rds")
```

